Raft Distributed Consensus Algorithm over WebRTC (Web Real-Time Communication)

### 2015-08-23 research_class/paper.tex
paper.tex, data/: drop acmtog, data/plots.

% Based on: v2-acmtog-sample.tex, dated March 7 2012
%
% Compilation using 'acmtog.cls' - version 1.2 (March 2012), Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
\documentclass{acmtog} % V1.2

\usepackage{hyperref}

@@

%\acmVolume{VV}
%\acmNumber{N}
%\acmYear{YYYY}
%\acmMonth{Month}
%\acmArticleNum{XXX}
%\acmdoi{10.1145/XXXXXXX.YYYYYYY}

\acmVolume{28}
\acmNumber{4}
\acmYear{2009}
\acmMonth{September}
\acmArticleNum{106}
\doi{10.1145/1559755.1559763}

%ISSN
\issn{1234-56789}

% Copyright
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

@@

\markboth{J. D. Martin et al.}{Raft Distributed Consensus Algorithm over WebRTC (Web Real-Time Communication)}

\author{Joel D. Martin {\upshape and} David Levine
\affil{University of Texas at Arglington}
}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
    <ccs2012>
    <concept>
    <concept_id>10003752.10003809.10010172</concept_id>
    <concept_desc>Theory of computation~Distributed
    algorithms</concept_desc>
    <concept_significance>500</concept_significance>
    </concept>
    </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Distributed algorithms}

%
% End generated code
%

\keywords{Distributed systems, Real-time chat, Distributed consensus, WWW, World Wide Web, Browsers}

%\acmformat{Vitor F. Pamplona, Manuel M. Oliveira, Gladimir V. G. Baranoski,
%and Sean Fogarty. 2009. Photorealistic models for pupil light reflex and iridal pattern deformation.
%{\em ACM Trans. Graph.} 28, 4, Article 106 (September 2009), 10 pages.\\
%\doiline}

@@

%\begin{bottomstuff}
%Manuel M. Oliveira acknowledges a CNPq-Brazil fellowship (305613/2007-3). Gladimir V. G. Baranoski acknowledges a
%NSERC-Canada grant (238337). Microsoft Brazil provided additional support.
%Authors' addresses: Sean Fogarty, (Current address) NASA Ames Research Center, Moffett Field, California 94035.
%\end{bottomstuff}

@@

\begin{enumerate}
\item \emph{Consistent order secure chat}: a chat room system where all users see the same message order but messages never traverse through a central server even though the application code itself is delivered from a central web service.
\item \emph{Private multiuser document editing/whiteboard system}: Google Docs among other systems has demonstrated a powerful model for collaborative document creation. However, these systems all user central web services to coordinate shared state and handle browser-to-browser communication. Using a Raft over WebRTC could enable tools with similar functionality but without the central coordination and communication.
\item \emph{Browser based network games}: most network games have certain states that must be agreed upon by all nodes (e.g. avatar dead or alive, etc). This can be achieved with Raft over WebRTC without requiring any data to traverse a central server. In addition to reducing browser to server bandwidth requirements, this may also allow lower latency gaming if games instances are partitioned by latency locality.
\end{enumerate}

@@

The current Raft.js implementation is based on "Consensus", Ongaro and
implements most of the Raft algorithm including: leader election, log
replication, safety, membership changes, client interaction, and
read-only operation optimizations. Log compaction and full client
linearizability are not currently implemented (clients and client
transactions are not assigned unique IDs). These author intends to
implement these functions in the future, however these do not have
a significant impact on the ability to test Raft combined with WebRTC.

The Raft protocol description in "Consensus", Ongaro, has undergone
some revisions since the original draft paper (including incorporation
of suggestions by the author of this paper). These changes include
cluster membership simplifications, separating the concepts of
committing entries to the log and applying entries to the state
machine, etc. As part of the work for implementing Raft
over WebRTC, the Raft.js implementation was updated to reflect the
most recent Raft paper.

@@

@@

\begin{figure*}[Ht]
\centerline{\includegraphics[width=15cm]{raft_rtc_architecture}}
\caption{Raft over RTC Architecture}
  \label{fig:raft_rtc_architecture}
\end{figure*}

@@

\item Raft.js over WebRTC: combine the existing Raft.js implementation with the small JavaScript application above to demonstrate Raft over WebRTC. This step uses a statically configured three node cluster (three browser tabs). The PeerJS signaling server is polled to retrieve the list and other clients and the Raft cluster is started when the number of PeerJS clients reaches three.

\item Updated Raft.js: update the Raft.js implementation to match the latest description of the Raft protocol (as described in "Consensus", Ongaro dissertation). This includes the implementation of the simplified membership change algorithm (sequential changes) and the separation of the log commit and state machine apply concepts. [Additionally, the new Raft.js implementation always starts a new Raft cluster as a cluster with a single node and additional nodes are dynamically added to the cluster (as suggested in section 4.4 of the paper).]

\item Dynamic Raft.js over WebRTC: enhance the signaling server with support for sessions/channels to enable multiple separate Raft clusters. Start each cluster with a single node and automatically add and remove Raft nodes as they come and go.

@@

\item Sessionless/Datagram RPC messages: extend Raft RPC requests and response messages so that RPC responses have enough context to be handled separately from RPC requests. Specifically this enables Raft to support a datagram (sessionless) network transport.

\item Atomic state machine operations: implement a state machine update operation that succeeds against the specified key only if no other operation has been applied to a key since the last time the client read from that key.

\item Automatic client request forwarding: extend the client request RPC to also be sessionless/datagram based. Implement a asynchronous JavaScript function that determines which node is the leader and forwards client request RPCs to the leader. This function must also correlate responses to the client request and invoke the callback for the original request.

\item Chat application: add the ability to send a message to the chat log using the client request forwarding functionality with an atomic state machine operation. Show the current chat history based on the chat entries in the state machine.

@@

\begin{figure*}[Ht]
\centerline{\includegraphics[width=15cm]{raft_rtc_sequence}}
\caption{Raft over RTC Sequence Diagram}
  \label{fig:raft_rtc_sequence}
\end{figure*}

@@

@@

\begin{figure*}[Ht]
\centerline{\includegraphics[width=15cm]{browser_1}}
\caption{Raft over WebRTC Browser with One Node}
  \label{fig:browser_1}
\end{figure*}

@@

\begin{figure*}[Ht]
\centerline{\includegraphics[width=15cm]{browser_2}}
\caption{Raft over WebRTC Browser with Five Node}
  \label{fig:browser_2}
\end{figure*}

@@

\begin{tikzpicture}
    \begin{axis}[
            title=Initial Cluster Activation,
            xlabel={Nodes},
            ylabel={Time (seconds)},
            grid=major,
        ]
        \addplot table {data/data_cluster_up.dat};
    \end{axis}
\end{tikzpicture}

The following figure shows the amount of time that a stable cluster takes to
propagate, commit and apply a single state machine change to the state
machines on every cluster node.

\begin{tikzpicture}
    \begin{axis}[
            title=Message Propagation,
            xlabel={Nodes},
            ylabel={Time (ms)},
            grid=major,
        ]
        \addplot table {data/data_propagate.dat};
    \end{axis}
\end{tikzpicture}

@@

\begin{tikzpicture}
    \begin{axis}[
            title=Node Failure Recovery,
            xlabel={Nodes},
            ylabel={Time (seconds)},
            grid=major,
            legend entries={$Leader Down$,$Half Cluster Down$},
        ]
        \addplot table {data/data_kill_leader.dat};
        \addplot table {data/data_kill_half.dat};
    \end{axis}
\end{tikzpicture}

@@

\begin{acks}
I would like to thank my academic supervisor, David Levine, for his
excellent input and oversight of this project. I would also like to
thank Diego Ongaro for his amazing work on the Raft protocol and paper
which made implementing a Raft implementation a pleasure. There are
too many people involved in the creation and standardization of WebRTC
but I would like to thank them for their tireless work to make
browser-to-browser communication possible. Last but not least, I would
like to thank my family for putting up with the long nights and
weekends I spent on this project. Thanks!
\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format-Journals}
% TODO: references
\bibliography{acmtog-sample-bibfile}
                                % Sample .bib file with references that match those in
                                % the 'Specifications Document (V1.5)' as well containing
                                % 'legacy' bibs and bibs with 'alternate codings'.
                                % Gerry Murray - March 2012

%\received{September 2008}{March 2009}

\end{document}
% End of v2-acmtog-sample.tex (March 2012) - Gerry Murray, ACM

>>>>

\documentclass[9pt]{extarticle}

\usepackage{multicol}
\setlength{\columnsep}{0.8cm}

\usepackage{geometry}
\geometry{margin=2.0cm}

\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}

@@

\author{Joel D. Martin and David Levine
\thanks{University of Texas at Arglington}
}

@@

@@

\begin{multicols}{2}

@@

\begin{enumerate}
\item \emph{Consistent order secure chat}: a chat room system where all users see the same message order but messages never traverse through a central server even though the application code itself is delivered from a central web service. This application was implemented for this study.
\item \emph{Private multiuser document editing/whiteboard system}: Google Docs among other systems has demonstrated a powerful model for collaborative document creation. However, these systems all user central web services to coordinate shared state and handle browser-to-browser communication. Using a Raft over WebRTC could enable tools with similar functionality but without the central coordination and communication.
\item \emph{Browser based network games}: most network games have certain states that must be agreed upon by all nodes (e.g. avatar dead or alive, etc). This can be achieved with Raft over WebRTC without requiring any data to traverse a central server. In addition to reducing browser to server bandwidth requirements, this may also allow lower latency gaming if games instances are partitioned by latency locality.
\end{enumerate}

@@

The current Raft.js implementation is based on "Consensus", Ongaro and
implements most of the Raft algorithm including: leader election, log
replication, safety, membership changes, basic client interaction, and
read-only operation optimizations. Log compaction and full client
linearizability are not currently implemented (clients and client
transactions are not assigned unique IDs). Log compaction is important
for long-running production systems in order to limit memory growth
and allow new nodes to catch up their logs more efficiently. However
log compaction is not required for verifying overall system
correctness and does not significantly impact short-running test
configurations. Full client linearizability is important feature for
simplifying certain types of distributed applications, however the
implementation requires some significant additions to the Raft
implemenation (e.g. assigning client IDs and sequence IDs and tracking
timed-out clients). For the chat application, simply adding one atomic
check-and-set state machine command is sufficient to achieve
a linearized chat history. This does impose a slight additional cost
on the client because it must retry the atomic operations until they
succeed. The author intends to implement both log compaction and
full client linearizability in the future but, however these do not
have a significant impact on the ability to test Raft combined with
WebRTC.

The Raft protocol description in "Consensus", Ongaro, has undergone
some revisions since the original draft paper (including incorporation
of suggestions by the author of this paper). These changes include
cluster membership simplifications, separating the concepts of
committing entries to the log and applying entries to the state
machine, clearer client interaction, etc. As part of the work for
implementing Raft over WebRTC, the Raft.js implementation was updated
to reflect these changes to the Raft protocol definition.

@@

\end{multicols}

\begin{figure*}[h]
\centerline{\includegraphics[width=0.8\textwidth]{raft_rtc_architecture}}
\caption{Raft over RTC Architecture}
  \label{fig:raft_rtc_architecture}
\end{figure*}

%\begin{multicols}{2}
%\end{multicols}

\begin{figure*}[h]
\centerline{\includegraphics[width=0.8\textwidth]{raft_rtc_sequence}}
\caption{Raft over RTC Sequence Diagram}
  \label{fig:raft_rtc_sequence}
\end{figure*}

\begin{multicols}{2}

@@

@@

\item Raft.js over WebRTC: combine the existing Raft.js implementation with the small JavaScript application above to demonstrate Raft over WebRTC. This step uses a statically configured three node cluster (three browser tabs). The PeerJS signaling server is polled to retrieve the list of other clients and the Raft cluster is started when the number of PeerJS clients reaches three.

\item Update Raft.js protocol: update the Raft.js implementation to match the latest description of the Raft protocol (as described in "Consensus", Ongaro dissertation). This includes the implementation of the simplified membership change algorithm (sequential changes) and the separation of the log commit and state machine apply concepts. Additionally, the new Raft.js implementation always starts a new Raft cluster as a cluster with a single node and additional nodes are dynamically added to the cluster (as suggested in section 4.4 of the paper).

\item Dynamic Raft.js over WebRTC: enhance the signaling server with support for multiple sessions/channels to enable separate Raft clusters with the same signalling server. Start each cluster with a single node and automatically add and remove Raft nodes as they come and go.

@@

\item Sessionless/Datagram RPC messages: change Raft RPC requests and response messages so that RPC responses have enough context to be handled separately from RPC requests. Specifically this enables Raft to support datagram (sessionless) network transports like WebRTC.

\item Atomic state machine operations: implement a state machine update operation that only succeeds if no other operation has been applied to the same key since the last time this client read from that key.

\item Automatic client request forwarding: extend the client request RPC to also be sessionless/datagram based. Implement an asynchronous JavaScript function that determines which node is the leader and forwards client request RPCs to the leader. This function must also correlate responses to the client request and invoke the callback for the original request.

\item Chat application: add the ability to send a message to the chat log using the client request forwarding functionality together with atomic state machine operation. Show the current chat history based on the chat entries in the state machine.

@@

@@

\subsubsection{Client Request Implementation}

TODO: client request functionality

\subsection{Chat Application}

The chat application extends the basic Raft.js over WebRTC client. The
chat application provides a large text area containing the chat
history and a text entry field below it. When the user clicks on the
\emph{Send} button (or pressed enter), the text line is added to
a pending send queue and the text entry field is blanked. A function
to flush sends runs periodically (currently every 100ms) and does the
following:

\begin{enumerate}
\item If there is line of text that has been sent but not yet
    added (applied) to the state machine or there are no lines of text
    on the pending send queue, then the function does nothing.
\item A text line is removed from the head of the pending send queue
    and stored in \emph{curSend} to indicate that a send is in
    progress. The a client request is made using an atomic
    \emph{seqAppend} command to add the new text to the state machine.
    The \emph{seqAppend} operation attempts to append the text line to
    a list at the \emph{history} key in the state machine.
\item When the atomic \emph{seqAppend} operation completes, it may
    have failed (if another text line was appended first), in which
    case the flush function puts the failed text line back on the
    pending send queue. Whether the send was successful or failed, the
    \emph{curSend} variable is set to nil/null to indicate that no
    send is in progress (the next time the flush function runs it will
    attempt to send again).
\end{enumerate}

TODO: screenshot.

TODO: atomic seqAppend operation

TODO: longer delay for propagation in chat application because each
addition to chat history take several messages to fully propagate.

@@

\begin{center}
    \includegraphics[width=0.4\textwidth]{browser_1.png}
    \captionof{figure}{Raft over WebRTC Browser with One Node}
    \label{fig:browser_1}
\end{center}

@@

%\begin{figure*}[Ht]
%\centerline{\includegraphics[width=15cm]{browser_2}}
%\caption{Raft over WebRTC Browser with Five Node}
%  \label{fig:browser_2}
%\end{figure*}

\begin{center}
    \includegraphics[width=0.4\textwidth]{browser_2.png}
    \captionof{figure}{Raft over WebRTC Browser with Five Nodes}
    \label{fig:browser_2}
\end{center}

In addition, the full chat application was also tested manually. TODO:
more text.

TODO: image of chat app with one node

TODO: image of chat app five nodes

@@

The following figure shows the amount of time that it takes for
a stable cluster to recover (elect a new leader and remove the failed
server from the cluster) when the current leader fails.

@@

\end{multicols}

\begin{figure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et100={mark=x,draw=blue},
                    et100_fail={mark=x,draw=red},
                    et100_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{100ms election timer}
        \label{fig:A}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et500={mark=x,draw=blue},
                    et500_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{500ms election timer}
        \label{fig:B}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=65,
                grid=major,
                scatter/classes={%
                    et1000={mark=x,draw=blue},
                    et1000_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_cluster_up_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{1000ms election timer}
        \label{fig:C}
    \end{subfigure}
    \caption{Initial Cluster Activation}
    \label{fig:cluster_up}

%%% Next Row
\rule{\textwidth}{0.3pt}
\smallskip

    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5, xmax=23,
                grid=major,
                scatter/classes={%
                    kl={mark=x,draw=blue},
                    kl_fail={mark=x,draw=red},
                    kl_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{100ms election timer}
        \label{fig:D}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kl={mark=x,draw=blue},
                    kl_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{500ms election timer}
        \label{fig:E}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kl={mark=x,draw=blue},
                    kl_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_leader_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{1000ms election timer}
        \label{fig:F}
    \end{subfigure}
    \caption{Recovery from Leader Failure}
    \label{fig:E}

%%% Next Row
\rule{\textwidth}{0.3pt}
\smallskip

    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5, xmax=23,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_100.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_100_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{100ms election timer}
        \label{fig:G}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_500.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_500_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{500ms election timer}
        \label{fig:H}
    \end{subfigure}
    \begin{subfigure}[b]{0.30\textwidth}
        \centering
        \resizebox{\linewidth}{!}{
            \begin{tikzpicture}
            \begin{axis}[%
                height=1.5\textwidth, width=1.4\textwidth,
                xlabel={Nodes}, ylabel={Time (seconds)},
                xtick=data,
                ymin=0, ymax=3.5,
                grid=major,
                scatter/classes={%
                    kh={mark=x,draw=blue},
                    kh_average={mark=,draw=red}}]
            \addplot[scatter,only marks,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_1000.dat};
            \addplot[scatter,%
                scatter src=explicit symbolic]%
            table[meta=label] {data/data_kill_half_1000_averages.dat};
            \end{axis}
            \end{tikzpicture}
        }
        \caption{1000ms election timer}
        \label{fig:I}
    \end{subfigure}
    \caption{Recovery from Half Cluster Failure}
    \label{fig:G}
\end{figure}

\begin{multicols}{2}

The following figure shows the amount of time that a stable cluster takes to
propagate, commit and apply a single state machine command to the state
machines on every cluster node.

\begin{center}
    \begin{tikzpicture}
    \begin{axis}[%
        xlabel={Nodes}, ylabel={Time (seconds)},
        xtick=data,
        grid=major,
        scatter/classes={%
            propagate={mark=x,draw=blue},
            propagate_average={mark=,draw=red}}]
    \addplot[scatter,only marks,%
        scatter src=explicit symbolic]%
    table[meta=label] {data/data_propagate.dat};
    \addplot[scatter,%
        scatter src=explicit symbolic]%
    table[meta=label] {data/data_propagate_averages.dat};
    \end{axis}
    \end{tikzpicture}
\captionof{figure}{Propagate$,$ Commit$,$ and Apply Command}
\label{fig:propagate}
\end{center}

@@

\section*{Acknowledgements}
I would like to thank my academic supervisor, David Levine, for his
excellent input and oversight of this project. I would also like to
thank Diego Ongaro for his amazing work on the Raft protocol and paper
which made implementing a Raft implementation a pleasure. There are
too many people involved in the creation and standardization of WebRTC
but I would like to thank them for their tireless work to make
browser-to-browser communication possible. Last but not least, I would
like to thank my family for putting up with the long nights and
weekends I spent on this project. Thanks!

% TODO: references

\end{multicols}
\end{document}
